Towards a Machine Learning approach for Aggregate Computing

### 2021-11-27 paper-2021-woa-ac-ml.tex
Pass on abstract

\maketitle
\begin{abstract}
Artificial intelligence is here to stay. 
 In particular, in recent years, Machine Learning techniques have outperformed 
 hand-craft human solutions in different contexts, ranging from computer vision to videogames.
%
%Currently, though, an all-encompassing Machine learning algorithm does not exist, 
% so researchers explore the possibility to apply these techniques in different fields.

An interesting field for Machine Learning applications is the design of algorithms for Collective-Adaptive Systems (CASs). % Or do we want to use Cyber-Physical Swarm??
Researchers have to deal with a series of challenges including dealing with massive number of nodes, global-to-local behaviour mapping, and highly environment stochasticity.  Historically, researchers have favoured the use of self-organisation, by which a collective behaviour emerges from continuously local interaction between simple components.
A novel technique to specify self-organising behaviour in a composable and functional manner is Aggregate Computing.
 In particular, this paradigm enables collective program specification as a whole, 
 as it is based on a distributed data structure manipulation called \textit{computational field}.
 It is successfully used in different scenarios ranging from Swarm Robotics to Smart cities.
Even though it is promising for expressing high-level collective behaviour system-size independent,
 practically it is difficult to write programs that work in highly dynamic networks, 
 where nodes have to deal with continuous changes in terms of environment.

Therefore, with this work, we propose a solution that can be seen as the first step towards the integration between Aggregate Computing and Machine Learning, and, in particular Reinforcement Learning. 
 The goal is to improve adaptivity through low-level specifications but maintaining high-level program behaviours.

In this paper, we outline issues, opportunities and methodologies of this integration, presenting a case study where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%

>>>>

\maketitle
\begin{abstract}
Recent trends like pervasive, autonomic, and swarm computing
 promote 
 a vision of large-scale collections of relatively simple agents
 that act and coordinate 
 with no central coordinator
 to support distributed applications.
%
Such Collective Adaptive Systems (CAS)
 are hard to engineer,
 because mechanisms for developing global behaviour
 out of local behaviour and interaction must be devised,
 often leveraging natural inspiration.
%
In particular, in the field-based Aggregate Computing approach,
 behaviour is specified in terms of computational fields,
 namely maps from CAS entities to computational values evolving over time, and neighbour-based interaction.
%
However, devising efficient aggregate computing algorithms
 for different scenarios
 is not an easy task; in particular,
 a problem is determining the best local decisions based on the current context
 for fostering collective outcomes.
%
Therefore, in this work we investigate
 the possibility of using machine learning techniques
 to automatically devise local actions
 to improve existing aggregate algorithms.
%
In particular, we propose a Reinforcement Learning solution that can be seen as the first step towards the integration between Aggregate Computing and Machine Learning. 
% The goal is to improve adaptivity through low-level specifications but maintaining high-level program behaviours. 
Accordingly, we outline challenges and opportunities of this integration, and present preliminary results based on a case study where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%
